Supermartingales and Termination Analysis of Statistical PCF

### 2020-06-09 relativised-completeness.tex
Fix minor errors in incomplete completeness proof.

For 1, consider the following:
\begin{proof}
If $T$ is bounded by $b$ a.s.~(for $n$ constant), the statement is trivially true by taking $\epsilon$ to be constantly 1, and $Y_n = b - n$.
Otherwise, let $(T_n)_{n \in \omega}$ be defined recursively such that $T_0 = 0$, $T_{n+1} > T_n$ and $\mathbb P[T < T_{n+1}] \leq \frac 1 4 \mathbb P[T < T_n]$. \lo{Since $T$ is a stopping time, we can assume that $T_n \in \mathbb N$. 
Problem with $T_1$: we require $T_1 > 0$ which may contradict the requirement $\mathbb P[T < T_{1}] \leq \frac 1 4 \mathbb P[T < 0] = 0$ for certain $T$, such as those that satisfy $\mathbb P[T < 1] > 0$.
We could define $T_1$ as the least number satisfying $\mathbb P[T < T_1] \geq \frac{1}{4}$.}
We will then define a supermartingale $(Y_n)_{n \in \omega}$ and nonrandom sequence $(Y'_n)_{n \in \omega}$ such that $Y_n = 0$ iff $T < n$ \changed[lo]{a.s.}, $Y_n = Y'_n$ iff $T \geq n$ \changed[lo]{a.s.}, and $Y'_n \geq 2^k$ for $n \geq T_k$.

@@

The antitone function \lo{It is good practice not to begin a sentence with a maths expression / symbol.} $\epsilon$ is defined piecewise and recursively in such a way as to force all the necessary constraints to hold:
\begin{align*}
    \text{for }x \in [0,1),\ &\epsilon(x) = 1 \\
    \text{for }x \in [2^k,2^{k+1}),\ &\epsilon(x) = \max \left (\epsilon(2^{k-1}), \frac{2^k}{T_{k+1}-T_k} \right ).
\end{align*}
This ensures that
\begin{itemize}
\item $\epsilon$ is weakly decreasing. 
\item $\epsilon$ is bounded below by 0.
\item For $x \geq 2^k$, $\epsilon(x) \leq \frac{2^k}{T_{k+1}-T_k}$. 
\lo{This would imply $\frac{2^k}{T_{k+1}-T_k} \geq \epsilon(2^{k-1})$. Why?}
\end{itemize}

@@

The fact $Y'_{T_k} \geq 2^{k+1}$ is proven by induction on $k$. For base case, $2 \geq 2$, as required. For the inductive case $k+1$, we first do another induction to prove that $Y'_n \geq 2^k$ for all $T_k \leq n \leq T_{k+1}$. The base case of this inner induction follows from the outer induction hypothesis a fortiori. take the greatest $k$ such that $n > T_k$. By the induction hypothesis, $Y'_{T_k} \geq 2^k$.
\begin{align*}
    Y'_n & = Y'_{T_k} \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} - \sum_{m=T_k}^{n-1} \epsilon(Y'_m) \frac{\mathbb P[T > m+1]}{\mathbb P[T > n]} \\
    & \geq \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (Y'_{T_k} - \sum_{m=T_k}^{n-1} \epsilon(Y'_m)) \\
    & \geq \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (Y'_{T_k} - \sum_{m=T_k}^{n-1} \frac{2^k}{T_{k+1}-T_k}) \\
    & = \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (Y'_{T_k} - \frac{2^k (n - T_k)}{T_{k+1}-T_k}) \\
    & \geq \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (Y'_{T_k} - 2^k) \\
    & \geq \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (2^{k+1} - 2^k) \\
    & = \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} 2^k \\
    & \geq 2^k
\end{align*}
as required. Substituting $n = T_{k+1}$ and using the fact that $\frac{\mathbb P[T > T_k]}{\mathbb P[T > T_{k+1}]}$, the same reasoning gives $Y'_{T_{k+1}} \geq 2^{k+2}$, as required.

>>>>

For 1, consider the following:
\begin{proof}
If $T$ is bounded by $b$ a.s.~(for $n$ constant), the statement is trivially true by taking $\epsilon$ to be constantly 1, and $Y_n = b - n$.
Otherwise, let $(T_n)_{n \in \omega}$ be defined recursively such that $T_0 = 0$, $T_{n+1} > T_n$ and $\mathbb P[T > T_{n+1}] \leq \frac 1 4 \mathbb P[T > T_n]$. \lo{Since $T$ is a stopping time, we can assume that $T_n \in \mathbb N$. 
Problem with $T_1$: we require $T_1 > 0$ which may contradict the requirement $\mathbb P[T < T_{1}] \leq \frac 1 4 \mathbb P[T < 0] = 0$ for certain $T$, such as those that satisfy $\mathbb P[T < 1] > 0$.
We could define $T_1$ as the least number satisfying $\mathbb P[T < T_1] \geq \frac{1}{4}$.} \akr{Sorry, I just wrote it wrong. I assume this ($\mathbb P[T > T_n]$ instead of $\mathbb P[T < T_n]$) resolves your concern.}
We will then define a supermartingale $(Y_n)_{n \in \omega}$ and nonrandom sequence $(Y'_n)_{n \in \omega}$ such that $Y_n = 0$ iff $T < n$ \changed[lo]{a.s.}, $Y_n = Y'_n$ iff $T \geq n$ \changed[lo]{a.s.}, and $Y'_n \geq 2^k$ for $n \geq T_k$.

@@

The antitone function \lo{It is good practice not to begin a sentence with a maths expression / symbol.} $\epsilon$ is defined piecewise and recursively in such a way as to force all the necessary constraints to hold:
\begin{align*}
    \text{for }x \in [0,1),\ &\epsilon(x) = 1 \\
    \text{for }x \in [2^k,2^{k+1}),\ &\epsilon(x) = \min \left (\epsilon(2^{k-1}), \frac{2^k}{T_{k+1}-T_k} \right ).
\end{align*}
This ensures that
\begin{itemize}
\item $\epsilon$ is weakly decreasing. 
\item $\epsilon$ is bounded below by 0.
\item For $x \geq 2^k$, $\epsilon(x) \leq \frac{2^k}{T_{k+1}-T_k}$. 
\lo{This would imply $\frac{2^k}{T_{k+1}-T_k} \geq \epsilon(2^{k-1})$. Why?} \akr{Corrected: it should have been min, not max.}
\end{itemize}

@@

The fact $Y'_{T_k} \geq 2^{k+1}$ is proven by induction on $k$. For base case, $2 \geq 2$, as required. For the inductive case $k+1$, we first do another induction to prove that $Y'_n \geq 2^k$ for all $T_k \leq n \leq T_{k+1}$. The base case of this inner induction follows from the outer induction hypothesis a fortiori. take the greatest $k$ such that $n > T_k$. By the induction hypothesis, $Y'_{T_k} \geq 2^k$.
\begin{align*}
    Y'_n & = Y'_{T_k} \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} - \sum_{m=T_k}^{n-1} \epsilon(Y'_m) \frac{\mathbb P[T > m+1]}{\mathbb P[T > n]} \\
    & \geq \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (Y'_{T_k} - \sum_{m=T_k}^{n-1} \epsilon(Y'_m)) \\
    & \geq \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (Y'_{T_k} - \sum_{m=T_k}^{n-1} \frac{2^k}{T_{k+1}-T_k}) \\
    & = \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (Y'_{T_k} - \frac{2^k (n - T_k)}{T_{k+1}-T_k}) \\
    & \geq \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (Y'_{T_k} - 2^k) \\
    & \geq \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} (2^{k+1} - 2^k) \\
    & = \frac{\mathbb P[T > T_k]}{\mathbb P[T > n]} 2^k \\
    & \geq 2^k
\end{align*}
as required. Substituting $n = T_{k+1}$ and using the fact that $\frac{\mathbb P[T > T_k]}{\mathbb P[T > T_{k+1}]} > 4$, the same reasoning gives $Y'_{T_{k+1}} \geq 2^{k+2}$, as required.
